\section{Domain-specific languages}
A \textit{domain-specific language (DSL)},
as opposed to a general-purpose language like C or Java,
targets a very specific application purpose.
Examples from various areas of software development are:
\begin{itemize}
  \item \textit{Markdown} for text formatting, \autocite{MarkdownIntro}
  \item the \textit{Gradle Build Language} for build configuration, \autocite{GradleWritingBuildScripts} and
  \item the \textit{Spock specification language} for test cases. \autocite{SpockFrameworkDoc}
\end{itemize}

These three languages can be organized in three rough categories,
distinguished by the way they are implemented:
\begin{enumerate}
  \item Standalone DSLs such as Markdown,
  \item entirely `userspace' DSLs such as the Gradle Build Language, and
  \item as a middle ground, custom compilation DSLs such as the Spock specification language.
\end{enumerate}

\paragraph{Standalone DSLs}
Standalone DSLs are implemented by defining a wholly new
lexical and syntactical grammar and the language semantics.
One can then use a parser generator to obtain a parser for the language
and include it in their software.

The major disadvantages of this approach for a testing DSL are
\begin{itemize}
  \item that the language would have to be enormously complex
    in order to be able to replace the regular Groovy language
    \begin{itemize}
      \item or it would have to be embedded in strings in the test cases, and
    \end{itemize}
  \item that there would be no tooling support for the language
    that provides autocompletion, formatting, error checking, and other assistance
    like there is for widespread existing languages.
\end{itemize}

\paragraph{Userspace DSLs}
\textit{Userspace DSLs} have recently become popular.
They are implemented entirely using features of the host language.
In particular, some languages are explicitly designed to
support the creation of DSLs well, such as
Groovy \autocite[Chapter 3.15: Domain-Specific Languages]{GroovyDoc}
and, more recently,
Kotlin \autocite[Chapter: Type-Safe Builders]{KotlinDoc}.
Language features that can aid the creation of DSLs are
\autocite[Chapter 3.15: Domain-Specific Languages]{GroovyDoc}
\begin{itemize}
  \item optional parentheses around call arguments,
  \item operator overloading,
  \item closure delegates, and
  \item compilation customization for advanced DSLs of the last category.
\end{itemize}

\paragraph{Custom compilation DSLs}
\textit{Custom compilation DSLs} also adhere to the syntax of an existing language,
but they offer more flexibility, because they are not restricted to runtime language features.
For example, labels in the code cannot be given additional semantics at runtime by a library,
which is something that, as we will see, the Spock Framework needs to do.

But unlike a standalone language,
existing tools can still handle the code written in the custom compilation DSL,
because all of it is valid code in the language it builds upon ---
although some tools may be confused by the
unorthodox usage of language features and become less helpful.

The compile-time portion of the implementation of this kind of DSL
primarily employs AST transformations,
which we will learn about in Chapter~\ref{chap:Ast}.
Groovy has extensive support for custom compilation and these transformations,
\autocite[Chapter 3.15.6: Compilation customizers]{GroovyDoc}
enabling the Spock Framework to implement its testing DSL cleanly.
