\section{Syntax}
Like assertion blocks, interaction declaration blocks are also introduced by a labeled statement.
The label name may be either \textit{mock} or \textit{stub},
with no technical distinction being made between those two ---
although users are, of course, urgently reminded not to use them inappropriately.

The body of such a labeled statement
--- or, in the case of a block statement, each statement contained within it ---
may have the shape of a \textit{Mock-}, \textit{Stub-}, or \textit{CombinedInteractionDeclaration}
as shown in Figure~\ref{fig:InteractionBlockVariants}.
The multiplication and right shift operators are the same
that the Spock Framework uses for interaction declarations.

\begin{figure}
  {\color{teal}
    \begin{minted}{text}
MockInteractionDeclaration:
cardinality * call;
    \end{minted}
  }
  {\color{magenta}
    \begin{minted}{text}
StubInteractionDeclaration:
call >> returnValue;
    \end{minted}
  }
  {\color{violet}
    \begin{minted}{text}
CombinedInteractionDeclaration:
cardinality * call >> returnValue;
    \end{minted}
  }
  \caption{
    The source code structure of a Mock-, Stub-, or CombinedInteractionDeclaration
  }\label{fig:InteractionBlockVariants}
\end{figure}

Figure~\ref{fig:InteractionBlockAst} shows the corresponding AST structure
for all three variants, distinguished by color.
The terminal nodes we need to extract are
\begin{itemize}
  \item the \textit{mockObject},
  \item the \textit{args},
  \item the \textit{cardinality} (for mock interactions), and
  \item the \textit{returnValue} (for stub interactions).
\end{itemize}
The \textit{mockObject} and the \textit{args} always appear
in the same \code{CallExpression} \textit{call}
that is shown in Figure~\ref{fig:InteractionBlockCallAst},
although that call may be attached to three different sections of the AST.

\begin{figure}
  \centering \setminted{fontsize=\footnotesize}
  \forestset{
    qtree/.style={
      baseline,
      for tree={
        align=center,
        l=3cm,
      }}}
  \begin{forest}, baseline, qtree
    [ExpressionStatement
    [ BinaryExpression (operator: {\color{teal}\code{*}} / {\color{magenta}\code{>>}} / {\color{violet}\code{>>}}), edge label={node[midway,left]{\code{expression}}}
    [  {\color{teal}Expression \code{cardinality}}\\{\color{magenta}CallExpression \code{call}}\\{\color{violet}BinaryExpression (operator: \code{*})}, edge label={node[midway,left]{\code{left}}}
    [   {\color{violet}Expression \code{cardinality}}, edge=violet, edge label={node[midway,left]{\code{left}}}]
    [   {\color{violet}CallExpression \code{call}}, edge=violet, edge label={node[midway,right]{\code{right}}}]]
    [  {\color{teal}CallExpression \code{call}}\\{\color{magenta}Expression \code{returnValue}}, edge label={node[midway,right]{\code{right}}}]]]
  \end{forest}
  \caption{
    The AST of a Mock-, Stub-, or CombinedInteractionDeclaration,
    matching colors with their corresponding source code structures
    in Figure~\ref{fig:InteractionBlockVariants}.
    The \textit{calls} are collapsed to save space;
    their homogeneous subtrees are shown in detail in Figure~\ref{fig:InteractionBlockCallAst}.
  }\label{fig:InteractionBlockAst}
\end{figure}

\begin{figure}
  \centering \setminted{fontsize=\footnotesize}
  \begin{forest}
  for tree={l=1.6cm}
    [{CallExpression \code{call}}
    [ {Expression \code{mockObject}}, edge label={node[midway,left]{\code{callee}}}]
    [ {Expression \code{args}}, edge label={node[midway,right]{\code{arguments}}}]]
  \end{forest}
  \caption{
    The AST of a \textit{call} as it appears in multiple places
    in Figure~\ref{fig:InteractionBlockAst}.
  }\label{fig:InteractionBlockCallAst}
\end{figure}

In addition to these interaction declaration blocks,
there are also interaction verification blocks
that are used to verify that the declared calls on a mock
have actually occurred:
\begin{minted}{javascript}
verify: mockObject;
\end{minted}
This explicit verification command is required to be test runner-agnostic,
because we would need to know the test case structure
for automatic insertion of verification statements.
We could still allow their omission by implementing automatic insertion
in test runner presets at some point.
